Negative and Fractional Types

### 2012-06-27 rational_ext.tex
pointed sets

\newcommand{\fname}[1]{\ulcorner #1 \urcorner}
\newcommand{\fconame}[1]{\llcorner #1 \lrcorner}

@@

%%%%%%%%%%%%%%%%%
\subsection{The Powerset Monad}

In order to interpret fractional types, we first define a monad {{S}} over
the {{Pi}}-category of finite sets and bijections.

>>>>

\newcommand{\vecspace}[1]{\mathbb{V}_{#1}}
\renewcommand{\vec}[2]{\begin{bmatrix} #1 \\ #2 \end{bmatrix}}
\newcommand{\Fpx}[1]{\mathbb{F}_{#1}}
\newcommand{\modulusp}[1]{\left | #1 \right |}
\newcommand{\fname}[1]{\ulcorner #1 \urcorner}
\newcommand{\fconame}[1]{\llcorner #1 \lrcorner}

@@

%%%%%%%%%%%%%%%%%
\subsection{Pointed Sets}

In order to interpret fractional types, we first lift the values and
combinators of {{Pi}} to the world of \emph{pointed sets}. Intuitively:
\begin{itemize}
\item instead of interpreting a type {{b}} as a plain set of cardinality
  $\modulusp{b}$, we now interpret each type {{b}} as a pointed set $P_b$
  with a distinguished \emph{basepoint} $\bot_b$;
\item a combinator {{c : b1 <-> b2}} is interpreted as a function as before
  with the restriction that it must map the basepoint $\bot_{b_1}$ to the
  basepoint $\bot_{b_2}$.
\end{itemize}

Intuitively a pointed set may be seen as a vector space over the field with
one element. This view allows us to add the type {{1/b}} and interpret it as
the \emph{dual} of the vector space denoting {{b}}. Concretely the elements
of the pointed set denoted by {{1/b}} are \emph{linear functionals} $f : P_b
\rightarrow P_1$. Note that $P_1$ is the pointed set containing {{()}} and
$\bot_1$. The isomorphism between the types {{1}} and {{b * 1/b}} is now a
consequence of that fact that the pairing of a vector and a functional in the
dual space produces a scalar.

\newpage
